---
title: "Iris Dataset with Linear Regression"
output: html_document
---

```{r}
datasets::iris
```

```{r}
head(iris)
```


```{r}
tail(iris)
```


```{r}
str(iris)
```


```{r}
summary(iris)
```


```{r}
plot(iris)

```

## Finding Correlation between features using correlation matrix

```{r}
cr <-cor(iris[1:4])
cr
```
### Using corrplot to plot the correlation matrix  


```{r}
library(corrplot)
corrplot(cr)


```

### Strong correlation is observed between sepal width & sepal length and Petal width and Petal Length  


## Sepal Length and Sepal WIdth analysis  


```{r}
sepal_len <- iris$Sepal.Length
hist(sepal_len, main= "Histogram for Sepal Length", xlab = "Sepal Length", xlim = c(4,8), col = "red", freq = FALSE)
```
```{r}
sepal_width <- iris$Sepal.Width
hist(sepal_width, main = "Histogram for Sepal Width", xlab = "Sepal Width",xlim = c(2,4.5), col = "darkorchid", freq = FALSE)
```

```{r}
library(ggplot2)
qplot(sepal_width,sepal_len, data = iris, color = Species)

```

### Linear Regression using Least Squares Method to find the regression line.

### Finding Correlation using Least Squares with Sepal Length as predictor and Sepal width as target

```{r}
corr <- cor(sepal_width , sepal_len, method = "pearson")
corr

```

```{r}
plot(sepal_width~sepal_len, col = sepal_len)
```

```{r}
model1 <- lm(sepal_width~sepal_len)
model1
```

```{r}
plot(sepal_width~sepal_len, col = sepal_len)
abline(model1 , col = "red" , lwd =3)
```
## Petal Length and Petal Width analysis

```{r}
petal_len <- iris$Petal.Length
petal_wid <- iris$Petal.Width
head(petal_len)
head(petal_wid)
```

```{r}
hist(petal_len , main = "Histogram of Petal Length", xlab = "Petal Length" , col = "blue", freq = FALSE,)
```
```{r}
hist(petal_wid , main = "Histogram of Petal WIdth" , xlab ="Petal Width", freq = FALSE, col = "green")
```

### Linear regression using least squares method lenth as predictor and width as target

```{r}
corr2 <- cor(petal_wid , petal_len, method = "pearson")
corr2

```
### Since the value is nearing 1 it shows a strong relationship

```{r}
model2 <- lm(petal_wid~petal_len)
model2
```

```{r}
plot(petal_wid~petal_len, col = petal_len)
abline(model2 , col = "blue" , lwd=3)
```

### Predicting Values

```{r}
#Prediction of Sepal width with Sepal length given as 15
p1 <- predict(model1 , data.frame(sepal_len=15))
p1
```

### Predicted value is 2.49 

```{r}
#Prediction of Petal width with petal length as 25
p2 <- predict(model2 , data.frame(petal_len=25))
p2

```

### Predicted value is 10.03  




